---
title: "Study of census attributes in Bandundu"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="C://Users//ecd1u18/Dropbox")
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```
```{r include=FALSE}
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

packages <- c("rgdal", "maptools", "spatialEco", 'MASS','tidyr',
              "raster", "sp", "gpclib", "rgeos",
              'tmap', 'dplyr', 'gsubfn', 'knitr')
ipak(packages)


df_resident <- sf::st_read(dsn = "data/in/eHealth/DRC_micocensus_Final.gdb",
                  layer = "residents")
df_hh <- sf::st_read(dsn = "data/in/eHealth/DRC_micocensus_Final.gdb",
                           layer = "household_data")
df_migr <- sf::st_read(dsn = "data/in/eHealth/DRC_micocensus_Final.gdb",
                           layer = "migrants")
census_res_geo <- readOGR("data/in/eHealth/DRC_micocensus_Final.gdb", 
                      "residential_buildings")
census_clust <- readOGR("data/in/microcensusCluster", 
                        "Microcensus_cluster_polygons")
boundaries <- readOGR("data/in/boundaries", "boundaries_bandunduKinshasa")
```


## Understanding incomplete survey

### Section 1: Overview

We study here the attributes of census observation in Bandundu that are reported as either `residential` or `mixte`. Our first concern goes to status of each observation, namely if it is completed or not. To do so let's understand what is under variable
```{r}
# In Bandundu
cluster_names <- read.csv('data/in/cluster_names.csv', row.names = 1)
census_res_geo <- census_res_geo[census_res_geo$cluster_id %in% cluster_names$mez_id,]
# Residential or Mixte building type
census_res_geo <- census_res_geo[census_res_geo$building_type %in% c('Residential', 'Mixed'),]
# For all tables
df_resident <- droplevels(df_resident[df_resident$cluster_id %in% cluster_names$mez_id,])
building_name <- df_resident %>% distinct(building_id) %>% select(building_id) %>% unlist(use.names = F)
df_hh <- droplevels(df_hh[df_hh$building_id %in% building_name,])
df_migr <- droplevels(df_migr[df_migr$building_id %in% building_name,])

census_res <- census_res_geo@data
```
```{r}
kable(as.data.frame(table(census_res_geo$status,dnn='Status' ), responseName='Count'),caption = 'Survey status for each observation' )
```
```{r message=TRUE}
tm_shape(boundaries) +
  tm_fill()+
  tm_shape(census_res_geo) +
  tm_dots(size=0.1,   
          col='status', title = 'Survey status \n (all building types)', palette = 'PiYG') +
  tm_legend()+
  tm_layout(title = 'Survey status', main.title.position= 'left',
            legend.outside = T, legend.outside.position = 'right')
```

### Section 2: Sources of false incompleteness 
#### 1. False report in `Survey_status`
*Situation*: Here we have on observation reported as `incomplete` but with household information filled and "nothing to report" on the `Comment` section. 

*Solution*: We change it to `complete`. Concerns **1** observation.

```{r }
kable(census_res_geo@data %>% filter(status=='Incomplete', population>1) %>% select(cluster_id, building_type, number_of_household, population, comments,status  ))
#census_res =census_res %>% mutate(cluster_id=replace(cluster_id, status=='Incomplete' & population >0, 'drc_maindombe_0062'))
census_res =census_res %>% mutate(status=replace(status, status=='Incomplete' & population >0, 'Complete'))
```




#### 2. Irrelevant buildings
*Situation*: as shown in the `comments`, some buildings reported as incomplete are actually not building that should contains people. Example: buildings reported abandonned, toilets, non occupied etc.

*Solution*: we remove them based on the pattern in the `Comments` section. Concerns **94** observations.

```{r echo=TRUE}
pattern = c('on occup', 'Abandonn', 'vide', 'noccup',
            "existe plus",'En construction','Boutique','habit',
            'Douche', 'Wc', 'detrui', 'toilette', 'Non occup',
            'Cuisine', 'Abondonnee','Deplaement', 'occup','Inau')
census_res =census_res %>% filter(!(status=='Incomplete' & grepl(paste(pattern, collapse = "|"),comments)))

```

#### 3. Duplicate of completed observations
*Situation*: Sometimes when the surveyor came back to the household, he/she didn't reopen the previous incomplete form such that for the same building we got two ids and two surveys, one completed and one incompleted (the first one in time). To flag it we based our analysis on the GPS coordinates (5 digits accuracy for the longitude, 6 for the lattitude), two buildings having the same coordinates were considered as the same. 

*Solution*: we remove these incomplete observations. Concerns **9** observations
```{r}
df_incomp <- census_res_geo@data[census_res_geo$status == 'Incomplete',]
df_comp <- census_res_geo@data[census_res_geo$status == 'Complete',]

incomp = data.frame(longitude=df_incomp$longitude,
                    latitude=df_incomp$latitude,
                    cluster_incomp = df_incomp$cluster_id,
                    building_incomp=df_incomp$building_id,
                    comments_incomp= df_incomp$comments)
incomp$incomp = TRUE
comp = data.frame(longitude=df_comp$longitude,
                  latitude=df_comp$latitude,
                  cluster_comp=df_comp$cluster_id,
                  building_comp=df_comp$building_id,
                  comments_comp= df_comp$comments)
comp$comp = TRUE

a=full_join(comp,incomp)

id_false_incomp <- as.character(a %>%  filter(comp==TRUE & incomp==TRUE) %>% distinct(building_incomp) %>% select(building_incomp) %>% unlist(use.names = F))
census_res = census_res %>% filter(!building_id %in% id_false_incomp) 
```
> Caveat: I'm not sure that this procedure is correct since in the full data table, there are 272 entries that have non-unique GPS coordinates.

```{r}
census_res_geo = census_res_geo[census_res_geo$building_id %in% census_res$building_id,]
kable(as.data.frame(table(census_res_geo$status,dnn='Status' ), responseName='Count'),caption = 'Survey status for each observation after modifications' )
```
```{r message=TRUE}
tm_shape(boundaries) +
  tm_fill()+
  tm_shape(census_res_geo) +
  tm_dots(size=0.1,   
          col='status', title = 'Survey status \n (all building types)', palette = 'PiYG') +
  tm_legend()+
  tm_layout(title = 'Survey status', main.title.position= 'left',
            legend.outside = T, legend.outside.position = 'right')
```

### Section 3 Caracteristics of incomplete survey

#### Brief description
For the remaining incomplete observations: we have information about:
- the number of floors (all at 1 except one at 174 considered as irrelevant outlier)
- the number of household in the building (all 1 except one at 2)

but no more. We have no other information for imputation.

## Analysis of completed survey

### Additional buildings surveyed
when we look at the `household` dataset, we notice that 30 buildings have been surveyed without being in the `building` dataset. They were reported as `non-residential` (even though they are absent from the `building` dataset corresponding to the non-residential ones).

What does it mean: we have more observations and we need to merge them with the others. It's possible because they have a building_id and a GPS coordinates. But we lacked the `comments`, the `interviewer_name`and the `number_of_floors` that are the information reported only at the building level.



```{r}
# Spot additional obs in the household dataset 
#   (ie building id not in the completed obs of the building dataset)
building_id = census_res$building_id
hh_building_id =df_hh %>% distinct(building_id) %>% select(building_id) %>% unlist(use.names = F)
supp_building =setdiff(hh_building_id, building_id)
paste('Number of additional observations:', length(supp_building)) # 30 buildings that are not in the master data

# Extract this additional data
df_supp <- df_resident %>% filter(building_id %in% supp_building) %>% select(-c(age, migrant, gender, resident_id, resident_number))
# Create from it their corresponding building observations
# Unkown variables at building level
df_supp <-df_supp %>% mutate(building_number=NA, building_type='residential',
                             interviewer_name=NA,number_of_floors=NA,
                             status='Complete', comments=NA)
# Compute variable at building level
#Number of households
hh_nb_supp <- df_supp %>% group_by(building_id) %>% distinct(household_id) %>% summarise(calculated_no_household=n())
#Number of people
pop_supp <- df_supp %>% group_by(building_id) %>% summarise(population=n())
#Number of migrants
migr_supp <- df_migr %>% filter(building_id %in% supp_building) %>% group_by(building_id) %>% summarise(number_of_migrants=n())
df_supp <- df_supp %>% left_join(hh_nb_supp) %>% 
            left_join(migr_supp) %>% left_join(pop_supp) %>% select(-household_id) %>% 
            filter(!duplicated(building_id)) 
df_supp$number_of_household=df_supp$calculated_no_household

# Merge it to the already existing building dataset
coords_supp <- lapply(df_supp %>% select(longitude,latitude), function(x) as.numeric(as.character(x)))
coords <- data.frame(coordinates(census_res_geo))
colnames(coords) <- c('longitude', 'latitude')
coords <- rbind(coords, coords_supp)
census_res <- rbind(census_res,df_supp)
census_res_geo <-  SpatialPointsDataFrame(coords, census_res,proj4string =  CRS(proj4string(census_res_geo)))
paste('Final number of buildings surveyed:', nrow(census_res_geo@data))

```

### Flag incompleteness at household level
We know the number of households that are supposed to live in the building from the `building` dataset and we know the number of households that have actually have been interviewed from the `household` dataset. We can thus flag buildings where not every household has been interviewed, information of prime importation for imputation purposes.

*NB*: `calculated_no_household` and `number_of_household` differs only for one observation that is



```{r}
hh_nb <- df_resident %>% group_by(building_id) %>% distinct(household_id) %>% summarise(hh_obs=n())
hh_complete= left_join(census_res, hh_nb) %>% mutate(nb_hh_incomp = as.integer(as.character(calculated_no_household)) - hh_obs )
kable(hh_complete %>% filter(nb_hh_incomp!= 0) %>% select(-c(building_number,building_id,interviewer_name,number_of_household,longitude, latitude, number_of_migrants,status)))
paste('Potential missing household observation:', nrow(hh_complete %>% filter(nb_hh_incomp!= 0)))
```



